Intentional release of native species undermines ecological stability

### 2022-10-28 rmd/manuscript.Rmd
Update manuscript.Rmd
fig reference

As with the two-species model, we ran `r attr(list_param[[2]], "n_sim")` timesteps of community dynamics with `r max(list_sim[[2]]$n_rep)` values of release level $R$ ($0$ to $500$ with a constant interval) for each simulation scenario. In this simulation, we initialized the community by introducing five individuals of each species and repeated this seeding every ten timesteps to account for possible complex dynamics of a ten-species community. In addition, we defined a threshold density ($N_i = 0.01$) below which species are removed from the simulation (i.e., an absorbing condition) and are recorded as "extinct". The rest of the simulation procedure is identical to the two-species model. We summarized values of simulation parameters in **Table S**\@ref(tab:tab-para).

@@

**Time-series data.** We assembled time-series fish data at `r n_distinct(d0$site_id)` sites within `r n_distinct(d0$river)` protected watersheds of Hokkaido Island, Japan. The Hokkaido Research Organization leads a long-term monitoring program at these watersheds, and the data are published as annual reports [@salmonandfreshwaterfisheriesresearchinstitutehokkaidoresearchorganizationAnnualReportMonitoring2021]. The program began in 1963, but an effective, standardized sampling method has been implemented since 1999 (two-pass sampling with a combination of electrofishing and cast net). Most data were collected in summer with irregular interannual intervals (1- to 3-year intervals for most cases), and sampling efforts were quantified by sampling area (average: `r op(mean(df_area$area))` $\pm$ `r op(sd(df_area$area))` m^2^). We confined our analysis to the sites that meet the following criteria: (i) the observation span (from the first to the last year of observation) exceeds ten years, (ii) the number of observation years exceeds five years, and (iii) masu salmon is observed at least twice during the observation period. As a result, we used time-series data at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` watersheds from 1999 to 2019. The summed abundance of first and second passes was used in the following analysis. **Table S**\@ref(tab:tab-splist) summarizes observed species in these watersheds.

@@

```{=tex}
\begin{align}
\begin{split}
&\ln n_{s,t} = \xi_{1,s} + \xi_{2,s} \ln n'_{s,t-1} + \varepsilon_{s,t}^{\text{state}}\\
&\ln n'_{s,t-1} = \frac{\sum_{q=1}^Q (\tau^{q} \ln n_{s,t-q})}{\sum_{q=1}^Q \tau^{q}}
\end{split}
(\#eq:state)
\end{align}
```
where $\xi_{1,s}$ is the site-specific constant at site $s$, $\xi_{2,s}$ the parameter characterizing an autoregressive (AR) process, and $\varepsilon_{s,t}^{\text{state}}$ the process error that follows a normal distribution as $\varepsilon_{s,t}^{\text{state}} \sim \text{Normal}(0, \sigma_{\text{state},s}^2)$. The parameters $\xi_{1,s}$ and $\xi_{2,s}$ were drawn from a multivariate normal distribution as $\pmb{\xi} \sim \text{MVN}(\pmb{\theta_{\xi}}, \pmb{\Omega_{\xi}})$, where $\pmb{\xi}$ is the matrix of $\xi_{1,s}$ and $\xi_{2,s}$. Equation \@ref(eq:state) is structurally equivalent to a $Q$-order AR model; however, we expressed the part of AR parameters as a geometric-series $\tau^q$ to reduce the number of parameters (our model has two AR parameters $\xi_{2,s}$ and $\tau$ regardless of order $Q$). The parameter $\tau$ controls how fast the density influence attenuates towards the past [@takagiTimescaleDependencyHost2015]. In this study, we set $Q=3$ given the life span of the study species (typically 1 -- 3 years). We used median estimates of fish density $n_{s,t}$ to calculate the temporal CV, mean ($\mu$), and SD ($\sigma$) for each site. For a whole community, we summed the densities of enhanced and unenhanced species. We summarized the reconstructed community dynamics in **Figures S**\@ref(fig:fig-d1) -- **S**\@ref(fig:fig-d3).

@@

```{=tex}
\begin{align}
\begin{split}
R'_s = \eta_{w,s'} R_w~\text{where}~
\sum_{s'}^{S_w + 1} \eta_{w,s'} = 1
(\#eq:effr)
\end{split}
\end{align}
```
In Equation \@ref(eq:effr), $\eta_{w,s'}$ represents the effective proportion of total release affecting the community at site $s'$ within watershed $w$. The summation of $\eta_{w,s'}$ over $S_w+1$ is assumed to be unity because a non-negligible portion of release may escape into unsurveyed areas. Since the values of $\eta_{w,s'}$ are not available in our dataset, we estimated them through a stochastic search while fitting the regression model to the data within a Bayesian framework (see **model fitting**). Other site-level predictors include upstream watershed area (log-transformed), air temperature, precipitation, and forest land use. Urban and agricultural land use was omitted because of either a limited value range (**Figure S**\@ref(fig:fig-env)) or a strong correlation with forest land use (**Figure S**\@ref(fig:fig-cor)). The watershed-specific intercept was related to a watershed-level predictor as:

@@

**Model fitting.** We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned vague or weakly informative priors to parameters (**Table S**\@ref(tab:tab-prior)). Four Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. The total MCMC iterations ranged from `r min(df_mcmc$n_total_mcmc)` to `r max(df_mcmc$n_total_mcmc)`, in which MCMC samples were saved every `r unique(df_mcmc$n_thin)` steps for the calculation of posterior distributions after the initial `r max(df_mcmc$n_burn)` burn-in period. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached $<1.1$ [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]. Parameter estimates were summarized in **Tables S**\@ref(tab:tab-ssm-ar) -- **S**\@ref(tab:tab-reg3).

>>>>

As with the two-species model, we ran `r attr(list_param[[2]], "n_sim")` timesteps of community dynamics with `r max(list_sim[[2]]$n_rep)` values of release level $R$ ($0$ to $500$ with a constant interval) for each simulation scenario. In this simulation, we initialized the community by introducing five individuals of each species and repeated this seeding every ten timesteps to account for possible complex dynamics of a ten-species community. In addition, we defined a threshold density ($N_i = 0.01$) below which species are removed from the simulation (i.e., an absorbing condition) and are recorded as "extinct". The rest of the simulation procedure is identical to the two-species model. We summarized values of simulation parameters in **Table S**1.

@@

**Time-series data.** We assembled time-series fish data at `r n_distinct(d0$site_id)` sites within `r n_distinct(d0$river)` protected watersheds of Hokkaido Island, Japan. The Hokkaido Research Organization leads a long-term monitoring program at these watersheds, and the data are published as annual reports [@salmonandfreshwaterfisheriesresearchinstitutehokkaidoresearchorganizationAnnualReportMonitoring2021]. The program began in 1963, but an effective, standardized sampling method has been implemented since 1999 (two-pass sampling with a combination of electrofishing and cast net). Most data were collected in summer with irregular interannual intervals (1- to 3-year intervals for most cases), and sampling efforts were quantified by sampling area (average: `r op(mean(df_area$area))` $\pm$ `r op(sd(df_area$area))` m^2^). We confined our analysis to the sites that meet the following criteria: (i) the observation span (from the first to the last year of observation) exceeds ten years, (ii) the number of observation years exceeds five years, and (iii) masu salmon is observed at least twice during the observation period. As a result, we used time-series data at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` watersheds from 1999 to 2019. The summed abundance of first and second passes was used in the following analysis. **Table S**2 summarizes observed species in these watersheds.

@@

```{=tex}
\begin{align}
\begin{split}
&\ln n_{s,t} = \xi_{1,s} + \xi_{2,s} \ln n'_{s,t-1} + \varepsilon_{s,t}^{\text{state}}\\
&\ln n'_{s,t-1} = \frac{\sum_{q=1}^Q (\tau^{q} \ln n_{s,t-q})}{\sum_{q=1}^Q \tau^{q}}
\end{split}
(\#eq:state)
\end{align}
```
where $\xi_{1,s}$ is the site-specific constant at site $s$, $\xi_{2,s}$ the parameter characterizing an autoregressive (AR) process, and $\varepsilon_{s,t}^{\text{state}}$ the process error that follows a normal distribution as $\varepsilon_{s,t}^{\text{state}} \sim \text{Normal}(0, \sigma_{\text{state},s}^2)$. The parameters $\xi_{1,s}$ and $\xi_{2,s}$ were drawn from a multivariate normal distribution as $\pmb{\xi} \sim \text{MVN}(\pmb{\theta_{\xi}}, \pmb{\Omega_{\xi}})$, where $\pmb{\xi}$ is the matrix of $\xi_{1,s}$ and $\xi_{2,s}$. Equation \@ref(eq:state) is structurally equivalent to a $Q$-order AR model; however, we expressed the part of AR parameters as a geometric-series $\tau^q$ to reduce the number of parameters (our model has two AR parameters $\xi_{2,s}$ and $\tau$ regardless of order $Q$). The parameter $\tau$ controls how fast the density influence attenuates towards the past [@takagiTimescaleDependencyHost2015]. In this study, we set $Q=3$ given the life span of the study species (typically 1 -- 3 years). We used median estimates of fish density $n_{s,t}$ to calculate the temporal CV, mean ($\mu$), and SD ($\sigma$) for each site. For a whole community, we summed the densities of enhanced and unenhanced species. We summarized the reconstructed community dynamics in **Figures S**11 -- **S**13.

@@

```{=tex}
\begin{align}
\begin{split}
R'_s = \eta_{w,s'} R_w~\text{where}~
\sum_{s'}^{S_w + 1} \eta_{w,s'} = 1
(\#eq:effr)
\end{split}
\end{align}
```
In Equation \@ref(eq:effr), $\eta_{w,s'}$ represents the effective proportion of total release affecting the community at site $s'$ within watershed $w$. The summation of $\eta_{w,s'}$ over $S_w+1$ is assumed to be unity because a non-negligible portion of release may escape into unsurveyed areas. Since the values of $\eta_{w,s'}$ are not available in our dataset, we estimated them through a stochastic search while fitting the regression model to the data within a Bayesian framework (see **model fitting**). Other site-level predictors include upstream watershed area (log-transformed), air temperature, precipitation, and forest land use. Urban and agricultural land use was omitted because of either a limited value range (**Figure S**14) or a strong correlation with forest land use (**Figure S**15). The watershed-specific intercept was related to a watershed-level predictor as:

@@

**Model fitting.** We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned vague or weakly informative priors to parameters (**Table S**3). Four Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. The total MCMC iterations ranged from `r min(df_mcmc$n_total_mcmc)` to `r max(df_mcmc$n_total_mcmc)`, in which MCMC samples were saved every `r unique(df_mcmc$n_thin)` steps for the calculation of posterior distributions after the initial `r max(df_mcmc$n_burn)` burn-in period. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached $<1.1$ [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]. Parameter estimates were summarized in **Tables S**4 -- **S**8.
